---
title: "Plotting Distribution Fits"
author: "Tynan"
date: "2023-10-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Plots

Made this on 2023-05-02, but it needs more work.

Made several different iterations of the distribution plots. This is still a WIP. Need to add normal distribution, was less interested in that at first.

Each iteration added either a new distribution or refined the code a bit to make it more compact. 

Currently, the functions both generate a data set based on a common distribution, plot it, and then assess the fit of the data to the distribution. 

Need to pull in some real-world data to demonstrate the use case of comparing the fit to different distributions. 

Also, should add some stat fit tests to complement the plot fit test. 


## Versions of Distribution Plots

* This version can fit log-normal, Weibull, and Beta distributions. 

* The control flow determines plot settings. But it's a bit repetitive, could be more compact. 

* NOTE: different params are needed for each distribution. 

```{r qqplot_and_histogram_v1, fig.width=15, fig.height=6}
qqplot_and_histogram_v1 <- function(distribution_type, sample_size, params) {
  
  library(cowplot)
  library(glue)  
  
  df <-  if (distribution_type == 'rlnorm') {
              as.data.frame( rlnorm(sample_size))
    
            } else if(distribution_type == 'rweibull') {
              as.data.frame( rweibull(sample_size, shape = params[1], scale = params[2])) 
              
            } else if(distribution_type == 'rbeta') {
              as.data.frame( rbeta(sample_size, shape1 = params[1], shape2 = params[2])) 

            } else if(distribution_type == 'rpois') {
              as.data.frame( rpois(sample_size, lambda = params[1]))

            }

  df <- df %>% 
        mutate(distribution = as.character(distribution_type),
               x            = row_number()) %>%
        rename(y = 1) %>% 
        dplyr::select(distribution, x, y)

  p1 <- df %>% 
        ggplot(.) + 
          aes(x=y) + 
          if (distribution_type == 'rpois') { 
          geom_bar()
          } else { 
          geom_histogram(binwidth = if_else(distribution_type %in% c('rlnorm', 'rweibull'),  0.1, 0.01),
                         fill = 'darkgreen') 
          } 
  
    p1 <- p1 + labs(title = glue('Histogram of {distribution_type}'))
  
         if (distribution_type == 'rlnorm') {
          p2 <- df %>% 
                 ggplot() +
                 aes(sample = y) +            
                 stat_qq(distribution      = stats::qlnorm, color = 'darkgreen') +
                 stat_qq_line(distribution = stats::qlnorm, color = 'darkgreen')  +
                 labs(title = glue('QQ plot of {distribution_type}')) 
          
         } else if(distribution_type %in% c('rweibull')) {
          p2 <- df %>% 
                 ggplot() +
                 aes(sample = y) + 
                 stat_qq(distribution      = stats::qweibull, dparams = list(shape = params[1], scale = params[2]), color = 'darkgreen') +
                 stat_qq_line(distribution = stats::qweibull, dparams = list(shape = params[1], scale = params[2]), color = 'darkgreen') +
                 labs(title = glue('QQ plot of {distribution_type}'))
          
         } else if(distribution_type %in% c('rbeta')) {
            p2 <- df %>%
                 ggplot() +
                 aes(sample = y) +
                 stat_qq(distribution      = stats::qbeta, dparams = list(shape1 = params[1], shape2 = params[2]), color = 'darkgreen') +
                 stat_qq_line(distribution = stats::qbeta, dparams = list(shape1 = params[1], shape2 = params[2]), color = 'darkgreen') +
                 labs(title = glue('QQ plot of {distribution_type}'))

         } else if(distribution_type %in% c('rpois')) {
            p2 <- df %>%
                 ggplot() +
                 aes(sample = y) +
                 stat_qq(distribution      = stats::qpois, dparams = list(lambda = params[1]), color = 'darkgreen') +
                 stat_qq_line(distribution = stats::qpois, dparams = list(lambda = params[1]), color = 'darkgreen') +
                 labs(title = glue('QQ plot of {distribution_type}'))

         }
  
  cowplot::plot_grid(p1, p2)  

}

qqplot_and_histogram_v1('rlnorm', 1000, c(5, 5))
```

* Does not fit Poisson distribution, but combines the Weibull and Beta conditions into the same section. 

```{r qqplot_and_histogram_v2, fig.width=15, fig.height=6}
qqplot_and_histogram_v2 <- function(distribution_type, sample_size, params) {
  
  library(cowplot)
  library(glue)  
  
  df <-  if (distribution_type == 'rlnorm') {
              as.data.frame( rlnorm(sample_size))
    
            } else if(distribution_type == 'rweibull') {
              as.data.frame( rweibull(sample_size, shape = params[1], scale = params[2])) 
              
            } else if(distribution_type == 'rbeta') {
              as.data.frame( rbeta(sample_size, shape1 = params[1], shape2 = params[2])) 

            }

  df <- df %>% 
        mutate(distribution = as.character(distribution_type),
               x            = row_number()) %>%
        rename(y = 1) %>% 
        dplyr::select(distribution, x, y)
    
  p1 <- df %>% 
        ggplot(.) + 
          aes(x=y) + 
          geom_histogram(binwidth = if_else(distribution_type %in% c('rlnorm', 'rweibull'),  0.1, 0.01))  +
          labs(title = glue('Histogram of {distribution_type}'))
  
         if (distribution_type == 'rlnorm') {
          p2 <- df %>% 
                 ggplot() +
                 aes(sample = y) +            
                 stat_qq(distribution      = stats::qlnorm) +
                 stat_qq_line(distribution = stats::qlnorm)  +
                 labs(title = glue('QQ plot of {distribution_type}')) 
          
         } else if(distribution_type %in% c('rweibull', 'rbeta')) {
          p2 <- df %>% 
                 ggplot() +
                 aes(sample = y) + 
                 stat_qq(distribution = if(distribution_type == 'rweibull') {stats::qweibull} else {stats::qbeta},
                         dparams      = if(distribution_type == 'rweibull') {
                                                  list(shape  = params[1], scale = params[2])}
                                           else { list(shape1 = params[1], shape2 = params[2]) }
                                          ) + 
                 stat_qq_line(distribution = if(distribution_type == 'rweibull') {stats::qweibull} else {stats::qbeta},
                              dparams      = if(distribution_type == 'rweibull') {
                                                    list(shape  = params[1], scale = params[2])}
                                             else { list(shape1 = params[1], shape2 = params[2]) }
                                                ) +                   
                 labs(title = glue('QQ plot of {distribution_type}'))

         }
  
  cowplot::plot_grid(p1, p2)  
  

}

qqplot_and_histogram_v2('rbeta', 1000, c(1, 5))
```

* This version is similar to qqplot_and_histogram_v1 above, need to come back and see what changed.

```{r qqplot_and_histogram_v3, fig.width=15, fig.height=6}
qqplot_and_histogram_v3 <- function(distribution_type, sample_size, params) {
  
  library(cowplot)
  library(glue)  
  
  df <-  if (distribution_type == 'rlnorm') {
              as.data.frame( rlnorm(sample_size))
    
            } else if(distribution_type == 'rweibull') {
              as.data.frame( rweibull(sample_size, shape = params[1], scale = params[2])) 
              
            } else if(distribution_type == 'rbeta') {
              as.data.frame( rbeta(sample_size, shape1 = params[1], shape2 = params[2])) 

            } else if(distribution_type == 'rpois') {
              as.data.frame( rpois(sample_size, lambda = params[1]))

            }

  df <- df %>% 
        mutate(distribution = as.character(distribution_type),
               x            = row_number()) %>%
        rename(y = 1) %>% 
        dplyr::select(distribution, x, y)

  p1 <- df %>% 
        ggplot(.) + 
          aes(x=y) + 
          if (distribution_type == 'rpois') { 
          geom_bar()
          } else { 
          geom_histogram(binwidth = if_else(distribution_type %in% c('rlnorm', 'rweibull'),  0.1, 0.01),
                         fill = 'darkgreen') 
          } 
  
    p1 <- p1 + labs(title = glue('Histogram of {distribution_type}'))
  
         if (distribution_type == 'rlnorm') {
          stat_qq_setting      <- stat_qq(distribution      = stats::qlnorm, color = 'darkgreen') 
          stat_qq_line_setting <- stat_qq_line(distribution = stats::qlnorm, color = 'darkgreen')  
          
         } else if(distribution_type %in% c('rweibull')) {
          stat_qq_setting      <- stat_qq(distribution      = stats::qweibull, dparams = list(shape = params[1], scale = params[2]), color = 'darkgreen')
          stat_qq_line_setting <- stat_qq_line(distribution = stats::qweibull, dparams = list(shape = params[1], scale = params[2]), color = 'darkgreen')
          
         } else if(distribution_type %in% c('rbeta')) {
          stat_qq_setting      <- stat_qq(distribution      = stats::qbeta, dparams = list(shape1 = params[1], shape2 = params[2]), color = 'darkgreen') 
          stat_qq_line_setting <- stat_qq_line(distribution = stats::qbeta, dparams = list(shape1 = params[1], shape2 = params[2]), color = 'darkgreen') 

         } else if(distribution_type %in% c('rpois')) {
          stat_qq_setting      <- stat_qq(distribution      = stats::qpois, dparams = list(lambda = params[1]), color = 'darkgreen') 
          stat_qq_line_setting <- stat_qq_line(distribution = stats::qpois, dparams = list(lambda = params[1]), color = 'darkgreen') 

         }

  p2 <- df %>%
        ggplot() +
        aes(sample = y) +
        stat_qq_setting +
        stat_qq_line_setting +
        labs(title = glue('QQ plot of {distribution_type}'))    
  
  cowplot::plot_grid(p1, p2)  


}

qqplot_and_histogram_v3('rbeta', 1000, c(1, 5))
```

* Final version

* The control flow for how the example data.frame is made has been updated. 

* A list called 'dist_args' is created to pass the params to a much more compact plotting section.

```{r qqplot_and_histogram_FINAL, fig.width=15, fig.height=6}

# final version 
qqplot_and_histogram <- function(distribution_type, sample_size, params) {
  
  library(cowplot)
  library(glue)  
  
   if (distribution_type == 'rlnorm') {
      df        <- as.data.frame( rlnorm(sample_size))
      dist_args =  list(dist_func = stats::qlnorm,
                        dparams   = NULL)

    } else if(distribution_type == 'rweibull') {
      df        <- as.data.frame( rweibull(sample_size, shape = params[1], scale = params[2])) 
      dist_args =  list(dist_func = stats::qweibull,
                        dparams   = list(shape  = params[1], scale = params[2]))
      
    } else if(distribution_type == 'rbeta') {
      df        <- as.data.frame( rbeta(sample_size, shape1 = params[1], shape2 = params[2])) 
      dist_args =  list(dist_func = stats::qbeta,
                        dparams   = list(shape1 = params[1], shape2 = params[2]))              

    } else if(distribution_type == 'rpois') {
      df        <- as.data.frame( rpois(sample_size, lambda = params[1])) 
      dist_args =  list(dist_func = stats::qpois,
                        dparams   = list(lambda = params[1]))

    }
  
  df <- df %>% 
        mutate(distribution = as.character(distribution_type),
               x            = row_number()) %>%
        rename(y = 1) %>% 
        dplyr::select(distribution, x, y)

  p1 <- df %>% 
        ggplot(.) + 
          aes(x=y) + 
          if (distribution_type == 'rpois') { 
          geom_bar(fill = 'darkgreen')
          } else { 
          geom_histogram(binwidth = if_else(distribution_type %in% c('rlnorm', 'rweibull'),  0.1, 0.01),
                         fill = 'darkgreen') 
          } 
  
  p1 <- p1 + labs(title = glue('Histogram of {distribution_type}'))
  
  p2 <- df %>%
        ggplot() +
        aes(sample = y) +
        stat_qq(distribution      = dist_args$dist_func, dparams = dist_args$dparams, color = 'darkgreen') +
        stat_qq_line(distribution = dist_args$dist_func, dparams = dist_args$dparams, color = 'darkgreen') +
        labs(title = glue('QQ plot of {distribution_type}'))    
  
  cowplot::plot_grid(p1, p2)  

}

qqplot_and_histogram('rlnorm', 1000)
```

## Final Version

* The final version of qqplot_and_histogram() is used when looping over different distributions.

```{r misc, fig.width=15, fig.height=6}
params_list <- list(c('rlnorm'),
                    c('rweibull', 1.2, 1.5),
                    c('rbeta',  1, 5),
                    c('rpois', 5))

for (i in seq_along( params_list)) {
  
  dist   <- params_list[[i]][1]
  params <- as.numeric(params_list[[i]][2:3])
  
  print(qqplot_and_histogram(dist, 1000, params))
  
}
```

## Preliminary Individul Distribution Plots 

```{r lognormal_hist}
y_dlnorm <- as.data.frame(rlnorm(10000)) %>% 
              mutate(x = row_number()) %>% 
              rename(y = 1)

y_dlnorm %>% 
ggplot(.) + 
  aes(x=y) + 
  geom_histogram(binwidth=0.1)
```

```{r logormal_qq_plot}
y_dlnorm %>% 
ggplot(., aes(sample = y)) +
   stat_qq(distribution = stats::qlnorm) +
   stat_qq_line(distribution = stats::qlnorm)
```

```{r weibull_hist}
weibull <- as.data.frame(rweibull(10000, shape = 1.2, scale = 1.5)) %>% 
              mutate(x = row_number()) %>% 
              rename(y = 1)

weibull %>% 
ggplot(.) + 
  aes(x=y) + 
  geom_histogram(binwidth=0.1)
```

```{r weibull_qq_plot}
weibull %>% 
ggplot() + 
   aes(sample = y) + 
   stat_qq(distribution      = stats::qweibull, dparams = list(shape = 1.2, scale = 1.5)) +
   stat_qq_line(distribution = stats::qweibull, dparams = list(shape = 1.2, scale = 1.5)) 
```

```{r weibull_hist_2}
weibull_2 <- as.data.frame(rweibull(1000, shape = 1.5, scale = 1)) %>% 
                mutate(x = row_number()) %>% 
                rename(y = 1)

weibull_2 %>% 
    ggplot() + 
    aes(x = y) + 
    geom_histogram(binwidth = 0.1)
```

```{r weibull_2}
weibull_2 %>% 
  ggplot() + 
  aes(sample = y) + 
  stat_qq(distribution      = stats::qweibull, dparams = list(shape = 1.5, scale = 1)) + 
  stat_qq_line(distribution = stats::qweibull, dparams = list(shape = 1.5, scale = 1)) 
```

```{r beta_hist}
beta <- as.data.frame(rbeta(10000, shape1 = 1, shape2 = 5)) %>% 
              mutate(x = row_number()) %>% 
              rename(y = 1)

beta %>% 
    ggplot() + 
    aes(x = y) + 
    geom_histogram(binwidth = 0.01)
```

```{r beta_plot}
beta %>% 
ggplot() + 
  aes(sample = y) + 
  stat_qq(distribution      = stats::qbeta, dparams = list(shape1 = 1, shape2 = 5)) + 
  stat_qq_line(distribution = stats::qbeta, dparams = list(shape1 = 1, shape2 = 5)) 
```

```{r poisson_hist}
poisson <- as.data.frame(rpois(1000, lambda = 5))  %>% 
              mutate(x = row_number()) %>% 
              rename(y = 1)

poisson %>% 
    ggplot() + 
    aes(x = y) + 
    geom_bar()
```

```{r poisson_plot}
poisson %>% 
  ggplot() + 
    aes(sample = y) + 
    stat_qq(distribution      = stats::qpois, dparams = list(lambda = 5)) + 
    stat_qq_line(distribution = stats::qpois, dparams = list(lambda = 5)) 
```

## Misc Plots 

* These are other distributions plots, could incorporate them into the main function, but need to get the qq line plot set-up. 

* This is for Student t Distribution. 

```{r t_dist}
data <- data.frame(x=rt(10000, df=7))

data %>% 
ggplot(aes(x=x)) +
  geom_histogram(aes(y = ..density..)) + 
  stat_function(fun       = dnorm, 
                linewidth = 1, 
                color     = 'gray', 
                args      = list()
                ) +
  labs(title="Histogram of interest rate changes") + 
  theme_bw()
```

* For The F Distribution

```{r t_dist_2}
data <- data.frame(x=rf(10000, df1=7, df2=120))

data %>% 
ggplot(aes(x=x)) +
  geom_histogram(aes(y = ..density..)) + 
  stat_function(fun       = dlnorm, 
                linewidth = 1, 
                color     = 'gray', 
                args      = list()) +
  labs(title="Histogram of interest rate changes",
       subtitle = "An example of the F Distribution") + 
  theme_bw()
```

## Chi Square Distribution Plot

```{r}
oestat = function(o, e){
  sum( (e-o)^2/e )
}

set.seed(1)
B = 10000
# here we pick an arbitrary length / not the same as for Celegans
n = 2847

expected = rep(n/4 ,4)
oenull = replicate(B, oestat(e=expected, o=rmultinom(1,size = n, prob = rep(1/4,4))))
```

```{r}
ggplot(data.frame(null_stats = oenull)) +
  geom_histogram(aes(x = null_stats), bins = 100, boundary=0)
```

```{r }
 ggplot(data.frame(stat = oenull), aes(sample = stat)) +
   stat_qq(distribution = stats::qchisq, dparams = list(df = 3)) +
   stat_qq_line(distribution = stats::qchisq, dparams = list(df = 3)) 
```
